---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# UKpopData

<!-- badges: start -->
<!-- badges: end -->

## About

Trying to gather population data by age and year historical and projections to use in regression models for driving test demand forecasts.


## Installation

to install from GitHub

`devtools::install_github("https://github.com/MarkNichollsDVSA/UKpopData")`





## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(UKpopData)
get_pop_data()
```



## Updating the data

Download the files as listed below and copy them into the `/inst/extdata/` folder


## Obtaining Historical data

It appears to be a static spreadsheet which changes name every time is it updated

Current release date: 25 June 2021  

Next Release due: September 2022

- Start website: [www.ons.gov.uk](https://www.ons.gov.uk) 
- Main menu: People, population and community Dropdown option: Population and migration
- In this section option: Population estimates Scroll down to: Datasets related to population estimates
- Select: Estimates of the population for the UK, England and Wales, Scotland and Northern Ireland
- Expand section: Mid-2001 to mid-2020 detailed time series edition of this dataset
- Supporting files you may find useful: UK population estimates, 1838 to 2020 (XLSX, 2.0 MB)

Direct link to current file: [UK population estimates, 1838 to 2020 (XLSX, 2.0 MB)](https://www.ons.gov.uk/file?uri=/peoplepopulationandcommunity/populationandmigration/populationestimates/datasets/populationestimatesforukenglandandwalesscotlandandnorthernireland/mid2001tomid2020detailedtimeseries/ukpopulationestimates18382020.xlsx)

## Obtaining Projected Data

From a tool from the ONS to download customised data

- Start website: [www.nomisweb.co.uk](https://www.nomisweb.co.uk)
- Data Downloads section: Query Data
- Select Datasets by source: Population Estimates/Projections
- Opened folder: National Population projections by single age of year


Make the following selections:

- Geography: Great Britain
- Ages: un-check all "Labour Market categories" and check only 3rd box in quick select section for Individual ages and 5 year age bands section, thereby selecting only ages 
- Select Years: Future years on from Historic data (2021 upwards to 2027, in this case )
- Select gender: Total, Males and Females
- Format selection: Microsoft Excel (.xlsx)
- Layout: projected year
- Other options: One table per worksheet checked

Wait,  then download data

Due to the nature of the web-application there is no static link to the current file

## Historical Data structure

The downloaded file has many tabs

The required table is called "Population estimates for UK, by sex and single year of age, mid-1971 to mid-2020" and is currently Table 3 (also tab "Table 3")

There are 3 tables, arranged vertically in this order: All Persons, Male then Female. All 3 tables have column per year, starting with the most recent, and one row per age, starting with 0 years, ages 90+ are aggregated into one row.


## Projected Data Structure

The downloaded file will have 3 tabs, each tab has one table.

The tabs are named appropriately: Total, Male and Female

All 3 tables have one column per year, starting with the oldest, and one row per age, starting with 0 up to the maximum age, without any aggregation for 90+
